As part of a broad plan for estates management by the University of Edinburgh, several hundreds of operations staff from the University’s Information Services Group (ISG), previously scattered across multiple sites, were rehoused in 2016/17 in open-plan offices within a refurbished ‘60s office block known as Argyle House. To complement the open-plan spaces, the ground floor of the premises was completely remodelled to contain an assortment of large and small meeting rooms. 

In a separate initiative that had been building momentum throughout 2016, ISG established an Internet of Things network network based on LoRaWAN, a secure long range and low power radio communications technology using the unlicensed spectrum for message broadcast by small battery-powered sensor devices. It was decided to help test the network by carrying out a pilot project to collect information about occupancy of the meeting rooms and environmental conditions in the office spaces more generally. This information was intended to help address two questions of concern to the University’s building managers: (i) was the configuration of meeting spaces appropriate for the needs ISG staff, and (ii) were the environmental conditions, particularly in terms of heating and ventilation, satisfactor in the open plan offices?

The pilot project deployed multiple devices, that were networked using a combination of Bluetooth and LoRaWAN protocols, In one of the small meeting rooms, sensors were attached to chairs and the door to measure temperature, light level and movement. In addition, sensors for light level, temperature, atmospheric pressure and relative humidity were placed near desks across one wing of an open-plan area.  

Before installing the sensors, we carried out a Privacy Impact Assessment (PIA) in which we were careful to demonstrate that no personal data would be captured by the sensors.\footnote{PIAs are becoming a critical part of the engagement process as part of new data protection legislation such as the General Data Protection Regulation (GDPR).} 
 Information about the pilot was communicated in two ways: by sending an initial email announcement to all building users, and by attaching a notice explaining the project to the wall of the monitored meeting room. As we will discuss shortly, however, subsequent investigation cast into doubt whether these steps were adequate in providing building users with an appropriate level of insight into the monitoring work. 

%%% Local Variables: ***
%%% mode:latex ***
%%% TeX-master: "main.tex"  ***
%%% End: ***